---
title: "Sensitivity analysis"
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---


# Load packages and data
```{r}
source("utils.R")
library(circlize)
library(ComplexHeatmap)
library(microViz)
library(mediation)
library(survival)
library(survminer)
library(lmtest)
source("nepisods_trols_glm.R")
source("GEE_prevalence_score.R")

load("data/vOTUs.processed.RA.RData", verbose = T)
load("resources/dada2_gtdb_2020_07_13.RData", verbose = T)
load("resources/J45_diagnosis/J45_cox_cross_220705.rdata", verbose = T)
load("resources/X.RData", verbose = T)
gut_score <- readxl::read_excel("resources/gut_scores_1y.xlsx")
family_host <- read_tsv("resources/fig1.data4.tab")
core_vfc <- read_tsv("results/core_vfc.tsv") |>
  mutate(group = case_when(prevalence_cat == "[-Inf,0.02)" ~ "very rare",
                           prevalence_cat == "[0.02,0.2)" ~ "rare",
                           prevalence_cat == "[0.2,0.5)" ~ "common",
                           prevalence_cat == "[0.5, Inf)" ~ "core")) |> 
  dplyr::select(OTU, prevalence, group) |> 
  rename(famid = OTU) |> 
  unique() |> 
  arrange(desc(prevalence))
phenodata_all <- X
physeq
```

# Preproc virome
```{r}
# Remove J45 NAs (no follow-up 5 years)
physeqf <- subset_samples(physeq, !is.na(j45_5yr_ever))
physeqf <- subset_taxa(physeqf, taxa_sums(physeqf)>0)
physeqf

# filter to temperate
physeqft <- subset_taxa(physeqf, virulence == " 0")
physeqft

# agglomerate to VFCs
physeqft_fam <- physeqft |> 
  speedyseq::tax_glom("famid")
taxa_names(physeqft_fam) <- paste0("f_", str_trim(tax_df(physeqft_fam)$famid))
physeqft_fam
```

# Preproc 16S
```{r}
# Remove other timepoints, remove NA asthma
phy <- subset_samples(phy, sample_data(phy)$Time == "1y")
phy
# add sample data
samdat <- sample_data(phy) |> data.frame() |> rename(abcno = ABCNO) |>
  mutate(abcno = as.character(abcno)) |> 
  left_join(J45_cox_cross |> mutate(abcno = as.character(abcno)))
rownames(samdat) <- sample_names(phy)
sample_data(phy) <- sample_data(samdat)

# filter to abcno in virome
phy <- subset_samples(phy, abcno %in% sample_df(physeqf)$abcno)
sample_names(phy) <- sample_df(phy)$abcno

phy <- prune_taxa(taxa_sums(phy) > 0, phy)
phy

tax_table(phy) <- tax_df(phy) |> 
  mutate(Genus = str_replace_all(str_trim(as.character(Genus)), "-", "_")) |> 
  dplyr::select(-tax) |> 
  as.matrix()

# Agglomerate to genus
print("STEP: Agglomerate taxonomic level")
phy.g <- speedyseq::tax_glom(phy, "Genus")
taxa_names(phy.g) <- paste0("g_", tax_df(phy.g)$Genus)
phy.g
```

# sPLS
## Subset to temperate core
```{r}
keep <- prevalence_df(physeqft_fam) |> filter(prevalence >= 50) |> pull(tax)
physeq_selected <- subset_taxa(physeqft_fam, taxa_names(physeqft_fam) %in% keep) |> 
  transform_phy(transform = "compositional") |> 
   transform_phy(transform = "log")

physeq_selected
```

## Create data partitions for nested cross-validation
```{r}
data <- physeq_selected |>
  otu_table() |> t() |> data.frame() |> 
  mutate(Class = as.character(sample_df(physeq_selected)$j45_5yr_ever))
table(data$Class)
head(data)[,1:10]

set.seed(11)
k <- 5 # number of out-cv folds
f <- createFolds(data$Class, k = k, 
                 list = TRUE, 
                 returnTrain = TRUE)

training <- list(); testing <- list()
for (i in 1:k) {
  print(paste("Fold: ", i))
  training.fold <- data[f[[i]],]
  testing.fold <- data[-f[[i]],]
  
  print(table(training.fold$Class))
  print(table(testing.fold$Class))
  
  training[[i]] <- training.fold 
  testing[[i]] <- testing.fold
  print("#####################")
}


## check that all test folds from the out-CV, contain all kids
bind_rows(testing) |> rownames() |> length()
```

## Inner CV
Iterate over outer-folds and do a 10-repeat 10-fold CV 
```{r, fig.width=8, fig.height=4}
ctrl <- caret::trainControl(method="repeatedcv",
                            number=10,
                            repeats=10,
                            returnResamp = "all",
                            savePredictions = "all",
                            allowParallel = T,
                            verboseIter=TRUE)

registerDoParallel(cores = 3)
set.seed(111)

splsda_nestedcv <- list(); reps_auc_list <- list();
max_auc_list <- list(); test_auc_list <- list(); 
test_preds_list <- list()

for (i in 1:k) {
  list.keepX = c(seq(2,20,1),
               seq(20, floor(ncol(training[[i]])/3),20),
               seq(ceiling(ncol(training[[i]])/3), floor(ncol(training[[i]])/2),25),
               seq(ceiling(ncol(training[[i]])/2), ncol(training[[i]])-1,30),
               ncol(training[[i]])-1)
  list.keepX
  ncv <- train(as.numeric(Class == "1") ~ .,
                      data = training[[i]],
                      method = get_mixOmics_spls(),
                      preProc = c("center", "scale"),
                      metric = "Rsquared",
                      tuneGrid = expand.grid(ncomp = 1,
                                             keepX = list.keepX, 
                                             keepY = 1),
                      trControl = ctrl)

  ## Validation set predictions
  reps_auc <- ncv |> 
    get_best_predictions() |> 
    group_by(Rep) |> 
    summarize(auc = as.numeric(pROC::auc(obs, pred, direction = "<")))|> 
    arrange(desc(auc))

  ### get max (best model)
  max_auc <- reps_auc |> pull(auc) |> max()
  
  splsda_nestedcv[[i]] <- ncv
  reps_auc_list[[i]] <- reps_auc
  max_auc_list[[i]] <- max_auc
  
  ## Test set predictions
  TestPreds <- predict(ncv, testing[[i]])
  cvauc.test <- auc(predictor = TestPreds, testing[[i]]$Class, direction = "<")
  test_preds_list[[i]] <- TestPreds
  test_auc_list[[i]] <- cvauc.test
}
```

### N features selected by fold
```{r}
auc_ncv_plots <- list()
for (i in 1:k) {
 auc_ncv_plots[[i]] <- auc_components_plot(splsda_nestedcv[[i]] )
}

cowplot::plot_grid(auc_ncv_plots[[1]] + ggtitle("Fold-1"),
                   auc_ncv_plots[[2]] + ggtitle("Fold-2"),
                   auc_ncv_plots[[3]] + ggtitle("Fold-3"),
                   auc_ncv_plots[[4]] + ggtitle("Fold-4"),
                   auc_ncv_plots[[5]] + ggtitle("Fold-5"),
                   ncol=2)
```

## Outer CV
Run models with optimised parameters in the entire training partitions of outer folds
```{r}
ctrl2 <- caret::trainControl(method="none",
                            returnResamp = "all",
                            savePredictions = "all",
                            allowParallel = T,
                            verboseIter=TRUE)

set.seed(111)
out.fit.list <- list(); out.testpreds <- list();
out.auc <- list()
keepX <- c(2, 20, 8, 20, 12) # based on visual inspection on previous plots
for (i in 1:k) {
  print(keepX[[i]])
  fit <- train(as.numeric(Class == "1") ~ .,
                      data = training[[i]],
                      method = get_mixOmics_spls(),
                      preProc = c("center", "scale"),
                      metric = "Rsquared",
                      tuneGrid = expand.grid(ncomp = 1,
                                             keepX = keepX[[i]],
                                             keepY = 1),
                      trControl = ctrl2, fixX = c(keepX[[i]]))

  out.fit.list[[i]] <- fit
  testPreds <- predict(fit, testing[[i]])
  out.testpreds[[i]] <- testPreds
  out.auc[[i]] <- auc(predictor = testPreds, testing[[i]]$Class, direction = "<")
}

unlist(out.auc)
unlist(out.auc) |> mean()
unlist(out.auc) |> sd()
```
## save data
```{r}
save(out.auc, 
     out.testpreds,
     out.fit.list,
   file="results/nested_cv.RData")
```


# Loadings interpretation
```{r}
taxhost <- tax_df(physeqft) |>
  dplyr::select(class, famid, family, OTU, hostFamily, hostGenus, tax) |>
  left_join(make_mradat(physeqft) |>
              dplyr::select(tax, mra, prevalence)) |>
  mutate(prevalence = prevalence*100)
```

# SFIG4
```{r}
## biological understanding using a 10-fold CV on all the data yields 19 VFCs (original manuscript)
load("results/spls_vfc.RData", verbose = TRUE)
original_loadings <- loadings_df_vfc |> 
  mutate(cv = "10-fold repeated CV")
```

## Loadings Outer-CV
```{r}
lds.outer <- list()
for (i in 1:k) {
   loadings_df_vfc <- out.fit.list[[i]] |>
    get_loadings("CV", remove_empty = F) |>
    mutate(cv = paste0("Outer fold-", i)) |> 
    mutate(famid = var, tax = var) |>
    left_join(tax_df(physeqft_fam) |>
                dplyr::select(class, famid, family) |>
                mutate(famid = paste0("f_", str_trim(famid))) |> unique()) |> 
    left_join(family_host |> mutate(famid = paste0("f_", str_trim(famid))), by=c("famid")) |> 
    left_join(make_mradat(physeqft_fam) |> dplyr::select(tax, mra)) |> 
    left_join(prevalence_df(physeqft_fam) |> dplyr::select(tax, prevalence)) |> 
    arrange(desc(abs(loading)), desc(sd)) |> 
    head(out.fit.list[[i]]$bestTune$keepX) |> 
    dplyr::select(-var, -tax) |> 
    left_join(core_vfc |> dplyr::select(famid, group)) 
  lds.outer[[i]] <- loadings_df_vfc 
}

loadings.outer <- original_loadings |> 
  bind_rows(lds.outer)

ggplot(loadings.outer |> mutate(f = paste0(family, " (", host, ")")), 
       aes(fct_reorder(f, prevalence), 
           loading, ymin = loading - sd, 
           ymax = loading + sd, fill=host)) +
  geom_errorbar() +
  geom_bar(stat = "identity", color="black", lwd=.3) + 
  coord_flip() + ylab("sPLS loadings") + xlab("VFCs") + 
  theme_minimal(base_size=10) +
  ylim(-1, 0.5) + 
  geom_hline(yintercept=0, linetype=2) +
  scale_fill_igv() +
  theme(legend.position = "bottom") +
  guides(fill="none") +
  facet_wrap(~ cv, nrow=1)

ggsave("figures/SFIG4.pdf", device="pdf",
       width=8, height=4)
```

```{r}
lds.outer <- bind_rows(lds.outer)
head(lds.outer)

lds.outer |> 
  group_by(class, family) |> 
  summarise(n = n_distinct(cv)) |> 
  mutate(is_original_loading = ifelse(family %in% original_loadings$family, TRUE, FALSE)) |> 
  left_join(original_loadings |> select(family, loading, sd)) |> 
  arrange(desc(n), desc(loading))
```

## Collect performances
```{r}
df <- data.frame(validation_best = unlist(max_auc_list),
                  test = unlist(out.auc))
df
mean(df$validation_best)
mean(df$test)
```


# NCV Virome scores
```{r, fig.width=6, fig.height=3}
vir_score <- unlist(out.testpreds)

virome_pls_score_vfc <- data.frame(abcno = names(vir_score),
                               pls_score = vir_score) |>
  left_join(dplyr::select(sample_df(physeqft), abcno, j45_5yr_ever, asthma_mother))

# ZSCORE - normalize the PLS score (scale and center)
virome_pls_score_vfc <- virome_pls_score_vfc |>
  mutate(zscore_vir_pls_score = as.numeric(scale(pls_score))) |> 
  rename(vir_pls_score = pls_score)
```

## NCV Virome score vs. asthma
```{r,fig.width=4, fig.height=3}
glm(j45_5yr_ever ~ zscore_vir_pls_score,
    data=virome_pls_score_vfc, family="binomial") |>
  tidylog(exp=T, conf.int = T)
```

# 16S scores (Stokholm et al.)
```{r}
gut_scores <- sample_df(physeqf) |> 
  dplyr::select(abcno, j45_5yr_ever) |> 
  left_join(gut_score |>
              mutate(abcno = as.character(abcno))) |> 
  dplyr::rename(bac_pls_score = j45_cross_pls_1y_all) |> 
  mutate(zscore_bac_pls_score = as.numeric(scale(bac_pls_score)))
head(gut_scores)
```

## 16S scores vs. asthma
```{r}
glm(j45_5yr_ever ~ zscore_bac_pls_score,
    data=gut_scores, family="binomial") |>
  tidylog(exp=T, conf.int = T)
```

```{r}
copdat_scores <- virome_pls_score_vfc |> 
  left_join(gut_scores) |> 
  filter(!is.na(bac_pls_score))
head(copdat_scores)
write_tsv(virome_pls_score_vfc, file="results/virome_scores_ncv.tsv")
write_tsv(copdat_scores, file="results/virome_bac_scores_ncv.tsv")
```

## EDFIG8A
```{r, fig.height=3.5, fig.width=5}
copdat_scores |> 
  glm(j45_5yr_ever ~ zscore_vir_pls_score*asthma_mother, data=_, family=binomial("logit")) |> 
  tidylog(exp=T, conf.int=T)
```

```{r}
# asthma_mother == 1
fit.1 <- copdat_scores |> 
  filter(asthma_mother == 1) |> 
  glm(j45_5yr_ever ~ zscore_vir_pls_score, data=_, family=binomial("logit")) |> 
  tidylog(exp=T, conf.int=T)
fit.1

# asthma_mother == 0
fit.2 <- copdat_scores |> 
  filter(asthma_mother == 0) |> 
  glm(j45_5yr_ever ~ zscore_vir_pls_score, data=_, family=binomial("logit")) |> 
  tidylog(exp=T, conf.int=T)
fit.2
```

```{r}
# asthma_mother == 1
fit.1x <- copdat_scores |> 
  filter(asthma_mother == 1) |> 
  glm(j45_5yr_ever ~ zscore_bac_pls_score, data=_, family=binomial("logit")) |> 
  tidylog(exp=T, conf.int=T)
fit.1x

# asthma_mother == 0
fit.2x <- copdat_scores |> 
  filter(asthma_mother == 0) |> 
  glm(j45_5yr_ever ~ zscore_bac_pls_score, data=_, family=binomial("logit")) |> 
  tidylog(exp=T, conf.int=T)
fit.2x
```

```{r}
df <- fit.2 |> 
  mutate(outcome = "asthma",
         group = "Non-asthmatic mother",
         data = "Virome") |> 
  bind_rows(fit.1 |> 
              mutate(outcome = "asthma",
                     group = "Asthmatic mother",
                     data = "Virome")) |> 
  bind_rows(fit.2x |> 
              mutate(outcome = "asthma",
                     group = "Non-asthmatic mother",
                     data = "16S")) |> 
  bind_rows(fit.1x |> 
              mutate(outcome = "asthma",
                     group = "Asthmatic mother",
                     data = "16S")) |> 
  filter(term != "(Intercept)") |> 
  dplyr::select(outcome, estimate, conf.low, conf.high, p.value, group, data)

df |> 
  ggplot(aes(x=data, y=estimate, color=group)) +
    geom_hline(yintercept=1, linetype=2, lwd=.2) +
    geom_point(position=position_dodge(width=0.2)) +
    geom_linerange(aes(xend=data, ymin=conf.low, ymax=conf.high),
                 position=position_dodge(width=0.2)) +
    coord_flip() +
    xlab("") + ylab("OR") +
    scale_color_colorblind() +
    #guides(color="none") +
    theme_pubr(base_size=12)
ggsave("figures/EDFIG8A.pdf", device="pdf", dpi=300,
       width=3, height=2.5)
```

# EDFIG8B
## NCV virome
```{r}
mygee(copdat_scores, "zscore_vir_pls_score", 5) + theme_pubr(base_size=10)
ggsave("figures/EDFIG8Bvir.pdf", device="pdf", dpi=300,
       width=4, height=3.5)
```

# Mediation analysis
### step 1: Y \~ X
```{r}
model.1 <- copdat_scores |>
              glm(j45_5yr_ever ~ zscore_vir_pls_score, data=_,
                  family="binomial")
                  
model.1 |> tidylog(conf.int=T, exp=T)
```

### step 2: M ~ X
```{r}
model.M <- copdat_scores |>
  lm(zscore_bac_pls_score ~ zscore_vir_pls_score,
     data=_)

tidylog(model.M, conf.int = T)
```

### step 3: Y \~ X + M
```{r}
model.Y <- copdat_scores |>
  mutate(asthma = as.numeric(j45_5yr_ever)-1) |>
  lm(asthma ~ zscore_vir_pls_score + zscore_bac_pls_score,
               data = _)

model.Y |> tidylog(conf.int = T, exp=T)
```

### step 4: mediation
```{r, fig.width=3, fig.height=4}
set.seed(11)
mediation_results <- mediation::mediate(model.M, model.Y,
                   treat='zscore_vir_pls_score',
                   mediator='zscore_bac_pls_score',
                   control.value=0, treat.value =1,
                   boot=TRUE, sims=1000)
summary(mediation_results)
```

### EDFIG8C
```{r}
data.frame(term = factor(c("Causal\nmediation effect\n (Virome through 16S)", "Direct effect\n(Virome)", "Total effect\n(Virome + 16S)"),
                         levels=c("Total effect\n(Virome + 16S)", "Direct effect\n(Virome)", "Causal\nmediation effect\n (Virome through 16S)")),
           estimate = c(0.01132, 0.03831, 0.04963),
           conf.low = c(0.00115, 0.00481, 0.01833),
           conf.high = c(0.02, 0.07, 0.08),
           group = factor(c(rep("Average", 3)),
                          levels=c("Average"))) |> 
  ggplot(aes(x=term, y=estimate)) +
  geom_point(aes(color=group), position=position_dodge(width=0.2)) +
  geom_linerange(aes(x=term, ymin=conf.low, ymax=conf.high, color=group),
                 position=position_dodge(width=0.2)) +
  coord_flip() +
  geom_hline(yintercept=0, linetype=2) +
  scale_color_manual(values=c(color_map2, c("Average" = "black"))) +
  xlab("") + ylab("Estimate") +
  theme_pubr(base_size=10) +
  guides(color="none")

ggsave("NM-02/NMR1_figures/forest_mediation.pdf", width=4, height=2.5)  
```

## moderator: asthma-mother
### step 2: M \~ X
```{r}
model.M <- copdat_scores |>
  lm(zscore_bac_pls_score ~ zscore_vir_pls_score*asthma_mother,
     data=_)

tidylog(model.M, conf.int = T)
```

### step 3: Y \~ X + M
```{r}
model.Y <- copdat_scores |>
  mutate(asthma = as.numeric(j45_5yr_ever)-1) |>
  lm(asthma ~ zscore_vir_pls_score*asthma_mother + zscore_bac_pls_score*asthma_mother,
               data =_)

model.Y |> tidylog(conf.int = T, exp=T)
```

### step 4: mediation
```{r, fig.width=3, fig.height=4}
set.seed(11)
mediation_results_mod <- mediation::mediate(model.M, model.Y,
                   treat='zscore_vir_pls_score',
                   mediator='zscore_bac_pls_score',
                   boot=TRUE, sims=10)
test.modmed(mediation_results_mod,
            covariates.1 = list(asthma_mother = 1),
            covariates.2 = list(asthma_mother = 0),
            sims=1000)
```

# EDFIG8D
```{r}
surv_data <- copdat_scores |> 
  mutate(abcno = as.character(abcno)) |> 
  mutate(vir_score_group = as.numeric(gtools::quantcut(zscore_vir_pls_score, q = 2))-1,
         bac_score_group = as.numeric(gtools::quantcut(zscore_bac_pls_score, q = 2))-1) |> 
  mutate(signature_group = case_when(bac_score_group == 1 & vir_score_group == 1 ~ "bac-H & vir-H",
                                     bac_score_group == 1 & vir_score_group == 0 ~ "bac-H & vir-L",
                                     bac_score_group == 0 & vir_score_group == 0 ~ "bac-L & vir-L",
                                     bac_score_group == 0 & vir_score_group == 1 ~ "bac-L & vir-H"),
         siganture_group = as.factor(signature_group),
         signature_scale3g = case_when(signature_group == "bac-L & vir-L" ~ 1,
                                     signature_group %in% c("bac-H & vir-L", "bac-L & vir-H") ~ 2,
                                     signature_group == "bac-H & vir-H" ~ 3),
         signature_scale4g = case_when(signature_group == "bac-L & vir-L" ~ 1,
                                     signature_group == "bac-H & vir-L" ~ 2,
                                     signature_group == "bac-L & vir-H" ~ 3,
                                     signature_group == "bac-H & vir-H" ~ 4)) |> 
  dplyr::select(abcno, zscore_vir_pls_score, zscore_bac_pls_score, vir_score_group, bac_score_group, signature_group, signature_scale3g, signature_scale4g, j45_5yr_ever, asthma_mother) |> 
  left_join(phenodata_all |> mutate(abcno = as.character(abcno)) |> dplyr::select(abcno, eventage_j45_5yr, event_j45_5yr))
head(surv_data)
```

```{r}
table(surv_data$signature_group)
```

```{r}
km_fit <- survfit(Surv(eventage_j45_5yr, event_j45_5yr == 1) ~ signature_group, data=surv_data)
summary(km_fit)
pl <- ggsurvplot(km_fit, fun = "event", conf.int=F,
           #conf.int.style = "step",  # customize style of confidence intervals
           xlab = "Time in days",
           break.time.by = 365.25,
           ggtheme = theme_pubr(),
           #risk.table = "abs_pct",
           pval = T)
pl
pdf("figures/EDFIG8D.pdf",
        width=5, height=5)
print(pl, newpage=F)
dev.off()
```

## Cox regression
```{r}
cox_4g <- coxph(Surv(eventage_j45_5yr, event_j45_5yr == 1) ~ signature_scale4g, data=surv_data)
broom::tidy(cox_4g, exp=T, conf.int=T)
```

```{r}
cox_3g <- coxph(Surv(eventage_j45_5yr, event_j45_5yr == 1) ~ signature_scale3g, data=surv_data)
broom::tidy(cox_3g, exp=T, conf.int=T)
```

# NCV virome score ~ TLR9
```{r}
# dominant model
virome_pls_score_vfc |> 
  mutate(abcno = as.character(abcno), j45_5yr_ever = as.factor(j45_5yr_ever)) |> 
  left_join(sample_df(physeqf) |> dplyr::select(abcno, tlr9_rs187084_G) |> mutate(abcno = as.character(abcno))) |> 
  mutate(score = as.factor(as.numeric(quantcut(zscore_vir_pls_score, q = 2))-1)) |>
  filter(!is.na(tlr9_rs187084_G)) |>
  mutate(tlr9 = case_when(tlr9_rs187084_G == 0 ~ 1,
                      tlr9_rs187084_G == 1 ~ 1,
                      tlr9_rs187084_G == 2 ~ 0)) |>
  lm(zscore_vir_pls_score ~ tlr9, data=_) |> 
  tidylog(conf.int=T)
```

# asthma ~ TLR9
```{r}
# dominant model
virome_pls_score_vfc |> 
  mutate(abcno = as.character(abcno), j45_5yr_ever = as.factor(j45_5yr_ever)) |> 
  left_join(sample_df(physeqf) |> dplyr::select(abcno, tlr9_rs187084_G) |> mutate(abcno = as.character(abcno))) |> 
  mutate(score = as.factor(as.numeric(quantcut(zscore_vir_pls_score, q = 2))-1)) |>
  filter(!is.na(tlr9_rs187084_G)) |>
  mutate(tlr9 = case_when(tlr9_rs187084_G == 0 ~ 1,
                        tlr9_rs187084_G == 1 ~ 1,
                        tlr9_rs187084_G == 2 ~ 0)) |>
  glm(j45_5yr_ever ~ tlr9, data=_, family="binomial") |>
  tidylog(exp=T, conf.int=T)
```

# asthma ~ NCV virome score + TLR9
```{r}
# dominant model
m0.dom <- virome_pls_score_vfc |> 
  mutate(abcno = as.character(abcno), j45_5yr_ever = as.factor(j45_5yr_ever)) |> 
  left_join(sample_df(physeqf) |> 
              dplyr::select(abcno, tlr9_rs187084_G) |> 
              mutate(abcno = as.character(abcno))) |> 
  mutate(score = as.factor(as.numeric(quantcut(zscore_vir_pls_score, q = 2))-1)) |>
  filter(!is.na(tlr9_rs187084_G)) |>
  mutate(tlr9 = case_when(tlr9_rs187084_G == 0 ~ 1,
                          tlr9_rs187084_G == 1 ~ 1,
                          tlr9_rs187084_G == 2 ~ 0)) |>
  glm(j45_5yr_ever ~ zscore_vir_pls_score + tlr9, data=_, family="binomial")

tidylog(m0.dom, exp=T, conf.int=T)
```

# asthma ~ NCV virome score + TLR9 + asthma*TLR9
```{r}
# dominant model
m1.dom <- virome_pls_score_vfc |> 
  mutate(abcno = as.character(abcno), j45_5yr_ever = as.factor(j45_5yr_ever)) |> 
  left_join(sample_df(physeqf) |> dplyr::select(abcno, tlr9_rs187084_G) |> mutate(abcno = as.character(abcno))) |>
  mutate(tlr9 = case_when(tlr9_rs187084_G == 0 ~ 1,
                          tlr9_rs187084_G == 1 ~ 1,
                          tlr9_rs187084_G == 2 ~ 0)) |> 
  mutate(score = as.factor(as.numeric(quantcut(zscore_vir_pls_score, q = 2))-1)) |>
  glm(j45_5yr_ever ~ zscore_vir_pls_score*tlr9, data=_, family="binomial") 

tidylog(m1.dom, exp=T, conf.int=T)
```

## LRT
```{r}
lrtest(m0.dom, m1.dom)
```

## EDFIG8F
```{r}
m1_aa <- virome_pls_score_vfc |> 
  mutate(abcno = as.character(abcno), j45_5yr_ever = as.factor(j45_5yr_ever)) |> 
  left_join(sample_df(physeqf) |> dplyr::select(abcno, tlr9_rs187084_G) |> mutate(abcno = as.character(abcno))) |>
  filter(tlr9_rs187084_G == 0) |> 
  glm(j45_5yr_ever ~ zscore_vir_pls_score, data=_, family="binomial") 

tidylog(m1_aa, exp=T, conf.int=T)

m1_ag <- virome_pls_score_vfc |> 
  mutate(abcno = as.character(abcno), j45_5yr_ever = as.factor(j45_5yr_ever)) |> 
  left_join(sample_df(physeqf) |> dplyr::select(abcno, tlr9_rs187084_G) |> mutate(abcno = as.character(abcno))) |>
  filter(tlr9_rs187084_G == 1) |> 
  glm(j45_5yr_ever ~ zscore_vir_pls_score, data=_, family="binomial") 

tidylog(m1_ag, exp=T, conf.int=T)

m1_gg <- virome_pls_score_vfc |> 
  mutate(abcno = as.character(abcno), j45_5yr_ever = as.factor(j45_5yr_ever)) |> 
  left_join(sample_df(physeqf) |> dplyr::select(abcno, tlr9_rs187084_G) |> mutate(abcno = as.character(abcno))) |>
  filter(tlr9_rs187084_G == 2) |> 
  glm(j45_5yr_ever ~ zscore_vir_pls_score, data=_, family="binomial") 

tidylog(m1_gg, exp=T, conf.int=T)
```

```{r}
data.frame(term = factor(c("AA", "AG", "GG"),
                         levels=c("AA", "AG", "GG")),
           estimate = c(1.4072178, 1.4072178, 0.6126870),
           conf.low = c(0.9679178, 1.0370917, 0.3189059),
           conf.high = c(1.9694150, 1.9217003, 1.0927735)) |> 
  ggplot(aes(x=term, y=estimate)) +
  geom_point(position=position_dodge(width=0.2)) +
  geom_linerange(aes(x=term, ymin=conf.low, ymax=conf.high),
                 position=position_dodge(width=0.2)) +
  xlab("") + ylab("Estimate") +
  geom_hline(yintercept=1, linetype=2, lwd=0.25) +
  theme_pubr(base_size = 12) +
  ylab("OR")
ggsave("figures/EDFIG8F.pdf", device="pdf", width=3, height=2.5)
```


# EDFIG8E
```{r}
m0 <- copdat_scores |>
  mutate(asthma = as.numeric(j45_5yr_ever)-1) |>
  glm(j45_5yr_ever ~ 1, data=_, family="binomial")

m1 <- copdat_scores |> 
  mutate(asthma = as.numeric(j45_5yr_ever)-1) |>
  glm(j45_5yr_ever ~ zscore_vir_pls_score, data=_, family="binomial")

m2 <- copdat_scores |> 
  mutate(asthma = as.numeric(j45_5yr_ever)-1) |>
  glm(j45_5yr_ever ~ zscore_vir_pls_score + zscore_bac_pls_score, data=_, family="binomial")

m3 <- copdat_scores |> 
  mutate(asthma = as.numeric(j45_5yr_ever)-1) |>
  glm(j45_5yr_ever ~ zscore_bac_pls_score, data=_, family="binomial")

lrtest(m1, m2) # vir vs. vir + bac
lrtest(m3, m2) # bac vs. vir + bac
```

```{r}
fit.vir <- glm(j45_5yr_ever ~ zscore_vir_pls_score, data=copdat_scores, family="binomial")
fit.bac <- glm(j45_5yr_ever ~ zscore_bac_pls_score, data=copdat_scores, family="binomial")
fit.vir_bac <- glm(j45_5yr_ever ~ zscore_vir_pls_score + zscore_bac_pls_score, data=copdat_scores, family="binomial")

r2.vir <- r2(fit.vir)
r2.bac <- r2(fit.bac)
r2.vir_bac <- r2(fit.vir_bac)

df <- data.frame(r2 = c(as.numeric(r2.vir), as.numeric(r2.bac), as.numeric(r2.vir_bac)), group=c("Virome", "16S", "Virome + 16S"))
ggplot(df, aes(x=group, y=r2*100, fill=group)) +
  geom_col() + 
  theme_pubr(base_size=10) +
  xlab("") +
  guides(fill="none")
ggsave("figures/EDFIG8E.pdf", device="pdf", width=2, height=2)  
```

# session info
```{r}
sessionInfo()
```