---
title: "sPLS classification"
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---


# Load packages and data
```{r}
source("utils.R")
library(circlize)
library(ComplexHeatmap)
library(microViz)
source("nepisods_trols_glm.R")

load("data/vOTUs.processed.RA.RData", verbose = T)
load("resources/dada2_gtdb_2020_07_13.RData", verbose = T)
load("resources/J45_diagnosis/J45_cox_cross_220705.rdata", verbose = T)
load("resources/X.RData", verbose = T)
gut_score <- readxl::read_excel("resources/gut_scores_1y.xlsx")
family_host <- read_tsv("resources/fig1.data4.tab")
core_vfc <- read_tsv("results/core_vfc.tsv") |>
  mutate(group = case_when(prevalence_cat == "[-Inf,0.02)" ~ "very rare",
                           prevalence_cat == "[0.02,0.2)" ~ "rare",
                           prevalence_cat == "[0.2,0.5)" ~ "common",
                           prevalence_cat == "[0.5, Inf)" ~ "core")) |> 
  dplyr::select(OTU, prevalence, group) |> 
  rename(famid = OTU) |> 
  unique() |> 
  arrange(desc(prevalence))

phenodata_all <- X
physeq
```

# Preproc virome
```{r}
# Remove J45 NAs (no follow-up 5 years)
physeqf <- subset_samples(physeq, !is.na(j45_5yr_ever))
physeqf <- subset_taxa(physeqf, taxa_sums(physeqf)>0)
physeqf

# filter to temperate
physeqft <- subset_taxa(physeqf, virulence == " 0")
physeqft

# agglomerate to VFCs
physeqft_fam <- physeqft |> 
  speedyseq::tax_glom("famid")
taxa_names(physeqft_fam) <- paste0("f_", str_trim(tax_df(physeqft_fam)$famid))
physeqft_fam
```

# Preproc 16S
```{r}
# Remove other timepoints, remove NA asthma
phy <- subset_samples(phy, sample_data(phy)$Time == "1y")
phy

# add sample data
samdat <- sample_data(phy) |> data.frame() |> rename(abcno = ABCNO) |>
  mutate(abcno = as.character(abcno)) |> 
  left_join(J45_cox_cross |> mutate(abcno = as.character(abcno)))
rownames(samdat) <- sample_names(phy)
sample_data(phy) <- sample_data(samdat)

# filter to abcno in virome
phy <- subset_samples(phy, abcno %in% sample_df(physeqf)$abcno)
sample_names(phy) <- sample_df(phy)$abcno

phy <- prune_taxa(taxa_sums(phy) > 0, phy)
phy

tax_table(phy) <- tax_df(phy) |> 
  mutate(Genus = str_replace_all(str_trim(as.character(Genus)), "-", "_")) |> 
  dplyr::select(-tax) |> 
  as.matrix()

# Agglomerate to genus
print("STEP: Agglomerate taxonomic level")
phy.g <- speedyseq::tax_glom(phy, "Genus")
taxa_names(phy.g) <- paste0("g_", tax_df(phy.g)$Genus)
phy.g
```

# sPLS
```{r}
keep <- prevalence_df(physeqft_fam) |> filter(prevalence >= 50) |> pull(tax)
physeq_selected <- subset_taxa(physeqft_fam, taxa_names(physeqft_fam) %in% keep) |> 
  transform_phy(transform = "compositional") |> 
   transform_phy(transform = "log")

physeq_selected
```

```{r}
data <- physeq_selected |>
  otu_table() |> t() |> data.frame() |> 
  mutate(Class = as.character(sample_df(physeq_selected)$j45_5yr_ever))


table(data$Class)
head(data)[,1:10]
```

## CV round 1
```{r, fig.width=5, fig.height=4}
ctrl <- caret::trainControl(method="repeatedcv",
                            number=10,
                            repeats=10,
                            returnResamp = "all",
                            savePredictions = "all",
                            allowParallel = T,
                            verboseIter=TRUE)

list.keepX = c(seq(2,20,1),
               seq(20, floor(ncol(data)/3),20),
               seq(ceiling(ncol(data)/3), floor(ncol(data)/2),25),
               seq(ceiling(ncol(data)/2), ncol(data)-1,30),
               ncol(data)-1)
list.keepX
doParallel::registerDoParallel(cores = 2)
set.seed(111)
splsda_model_relab_all_fam <- train(as.numeric(Class == "1") ~ .,
                      data = data,
                      method = get_mixOmics_spls(),
                      preProc = c("center", "scale"),
                      metric = "Rsquared",
                      tuneGrid = expand.grid(ncomp = 1,
                                             keepX = list.keepX,
                                             keepY = 1),
                      trControl = ctrl)


reps_auc <- splsda_model_relab_all_fam |>
  get_best_predictions() |>
  group_by(Rep) |>
  summarize(auc = as.numeric(pROC::auc(obs, pred, direction = "<")))|>
  arrange(desc(auc))

# print AUC for each repeat
reps_auc

# extract best repetition
bestRep <- splsda_model_relab_all_fam |>
  get_best_predictions() |> group_by(Rep) |>
  summarize(auc = as.numeric(pROC::auc(obs, pred, direction = "<"))) |>
  arrange(desc(auc)) |>
  (\(x) x[1, "Rep"])() |>
  as.character()

# extract predictions for calculating CV AUC from best repetition
savedPreds <- splsda_model_relab_all_fam |>
  get_best_predictions() |>
  filter(Rep == bestRep)
cvauc.train <- auc(obs ~ pred, direction = "<", data = savedPreds)
cvauc.train

# plot AUC for each number of features
vfc_auc_1round <- auc_components_plot(splsda_model_relab_all_fam)
vfc_auc_1round
ggsave("figures/SFIG2.pdf", device="pdf", dpi=300, height=4, width=6)
```

## CV round 2
```{r, fig.width=8, fig.height=4}
registerDoParallel(cores = 2)
set.seed(111)
splsda_model_relab_all_fam2 <- train(as.numeric(Class == "1") ~ .,
                      data = data,
                      method = get_mixOmics_spls(),
                      preProc = c("center", "scale"),
                      metric = "Rsquared",
                      tuneGrid = expand.grid(ncomp = 1, 
                                             keepX = seq(10, 25, 1), 
                                             keepY = 1),
                      trControl = ctrl)


reps_auc <- splsda_model_relab_all_fam2 |> 
  get_best_predictions() |> 
  group_by(Rep) |> 
  summarize(auc = as.numeric(pROC::auc(obs, pred)))|>
  arrange(desc(auc))
reps_auc

bestRep <- splsda_model_relab_all_fam2 |> 
  get_best_predictions() |> group_by(Rep) |> 
  summarize(auc = as.numeric(pROC::auc(obs, pred)))|> 
  arrange(desc(auc)) |>
  (\(x) x[1, "Rep"])() |> 
  as.character()

savedPreds <- splsda_model_relab_all_fam2 |> 
  get_best_predictions() |> 
  filter(Rep == bestRep)
cvauc.train <- auc(obs ~ pred, data = savedPreds)

auc_components_plot(splsda_model_relab_all_fam2)
```

## CV round 3
```{r, fig.width=8, fig.height=4}
registerDoParallel(cores = 3)
set.seed(111)
splsda_model_relab_all_fam3 <- train(as.numeric(Class == "1") ~ .,
                      data = data,
                      method = get_mixOmics_spls(),
                      preProc = c("center", "scale"),
                      metric = "Rsquared",
                      tuneGrid = expand.grid(ncomp = 1,
                                             keepX = 19, 
                                             keepY = 1),
                      trControl = ctrl, fixX = c(19))


reps_auc <- splsda_model_relab_all_fam3 |> 
  get_best_predictions() |> 
  group_by(Rep) |> 
  summarize(auc = as.numeric(pROC::auc(obs, pred, direction = "<")))|> 
  arrange(desc(auc))
reps_auc

bestRep <- splsda_model_relab_all_fam3 |> 
  get_best_predictions() |> group_by(Rep) |> 
  summarize(auc = as.numeric(pROC::auc(obs, pred, direction = "<"))) |>
  arrange(desc(auc)) |>
  (\(x) x[1, "Rep"])() |> 
  as.character()

savedPreds <- splsda_model_relab_all_fam3 |> 
  get_best_predictions() |>
  filter(Rep == bestRep)
cvauc.train <- auc(obs ~ pred, direction = "<", data = savedPreds)

splsda_model_relab_all_fam3$bestTune
```

## FIG3A
```{r, fig.width=4, fig.height=4}
df = data.frame(trainpreds = savedPreds$pred,
                class = ifelse(savedPreds$obs == 1, "Asthma", "Healthy"))

ggplot(df, aes(x=class, y=trainpreds)) +
  geom_point(aes(color=class, fill=class), pch=21, size=2.5, position=position_jitter(h=0,w=.1)) +
  geom_boxplot(aes(fill=class), outlier.size=-Inf, width=0.5/2, alpha=.6) +
  scale_fill_manual(values = color_map2) + scale_color_manual(values =  color_rev_map2) +
  guides(color="none", fill="none") +
  xlab("") +
  ylab("Predictions") +
  ggtitle(paste0("Repeated ",
                 splsda_model_relab_all_fam3$control$number, "-fold CV AUC = ", roundex(cvauc.train, 2))) +
  theme_minimal(base_size=12)

ggsave("figures/FIG3A.pdf", device="pdf", dpi=300, width=4, height=3.5)
```


## Loadings
```{r}
## Extract most abundant host
taxhost <- tax_df(physeqft) |>
  dplyr::select(class, famid, family, OTU, hostFamily, hostGenus, tax) |>
  left_join(make_mradat(physeqft) |>
              dplyr::select(tax, mra, prevalence)) |>
  mutate(prevalence = prevalence*100)
```

```{r}
loadings_df_vfc <- splsda_model_relab_all_fam3 |>
  get_loadings("CV", remove_empty = F) |>
  mutate(famid = var, tax = var) |>
  left_join(tax_df(physeqft_fam) |>
              dplyr::select(class, famid, family) |>
              mutate(famid = paste0("f_", str_trim(famid))) |> unique()) |> 
  left_join(family_host |> mutate(famid = paste0("f_", str_trim(famid))), by=c("famid")) |> 
  left_join(make_mradat(physeqft_fam) |> dplyr::select(tax, mra)) |> 
  left_join(prevalence_df(physeqft_fam) |> dplyr::select(tax, prevalence)) |> 
  arrange(desc(abs(loading)), desc(sd)) |> 
  head(splsda_model_relab_all_fam3$bestTune$keepX) |> 
  dplyr::select(-var, -tax) |> 
  left_join(core_vfc |> dplyr::select(famid, group))

ggplot(loadings_df_vfc |> mutate(f = paste0(family, " (", host, ")")), 
       aes(fct_reorder(f, prevalence), 
           loading, ymin = loading - sd, 
           ymax = loading + sd, fill=host)) +
  geom_errorbar() +
  geom_bar(stat = "identity", color="black", lwd=.3) + 
  coord_flip() + ylab("sPLS loadings") + xlab("VFCs") + 
  theme_minimal(base_size=12) +
  ylim(-0.6, 0.2) + 
  scale_fill_igv() +
  theme(legend.position = "bottom") +
  guides(fill="none")

ggsave("figures/FIG3B.pdf", device="pdf", dpi=300, height=4, width=6)
```

## most common predicted hosts across all 19 VFC
```{r}
taxhost |> 
  filter(paste0("f_", str_trim(famid)) %in% loadings_df_vfc$famid) |> 
  dplyr::select(famid, hostGenus) |> 
  filter(!is.na(hostGenus)) |> 
  group_by(hostGenus) |> 
  summarise(n = n()) |> 
  arrange(desc(n))
```

## chordDiagram
```{r}
df <- taxhost |> 
  mutate(famid = paste0("f_", str_trim(famid))) |> 
  filter(famid %in% loadings_df_vfc$famid) |> 
  dplyr::select(hostGenus, famid, family) |> 
  filter(!is.na(hostGenus)) |> 
  group_by(hostGenus, family, famid) |> 
  summarise(value = n()) |> 
  arrange(family)

# filter hosts with only one vOTU link
filt <- taxhost |> 
  mutate(famid = paste0("f_", str_trim(famid))) |> 
  filter(famid %in% loadings_df_vfc$famid) |> 
  dplyr::select(family, hostGenus, famid) |> 
  filter(!is.na(hostGenus)) |>
  group_by(hostGenus) |> 
  summarise(n = n()) |> 
  arrange(n) |> 
  filter(n > 1)

df <- df |>
  filter(hostGenus %in% filt$hostGenus)
```

```{r}
col.host <- c("Ruminococcus" = "#2596be",
              "Lactococcus" = "#6bd76b",
              "Bacteroides" = "#749b58", 
              "Flavonifractor" = "#802268",
              "Barnesiella" = "#f0e685",
              "Anaerostipes" = "#5050ff",
              "Blautia" = "#466983",
              "Faecalibacterium" = "#5db1dd",
              "Tyzzerella" = "#924822",
              "Collinsella" = "#ba6338",
              "Bacillus" = "#ce3d32",
              "Clostridium" = "#8cbb9c",
              "Coprococcus" = "#8cacbc",
              "Dorea" = "#bc9c8c",
              "Enterococcus" = "#bcb48c",
              "Lachnoclostridium" = "#9c8cbc",
              "Lactobacillus" = "#acbc8c",
              "Streptococcus" = "#886464",
              "Prevotella" = "#06768e",
              "Oscillobacter" = "#8e1e06",
              "Paenibacillus" = "#8e6206",
              "Roseburia" = "#768e06",
              "Herbaspirillum" = "#1e068e",
              "Parabacteroides" = "#ad97a9",
              "Erysipelatoclostridium" = "darkgreen",
              "Eubacterium" = "goldenrod3",
              "Clostridioides" = "peachpuff",
              "Oscillibacter" = "orchid3",
              "Haemophilus" = "khaki3")

col.vfc <- c("Irisviridae" = "#212121", #VOC1
             "Sidselviridae" = "#212121",
             "Hugoviridae" = "#212121",
             "Madsviridae" = "#212121",
             "Valdemarviridae" = "#212121",
             "Ludvigviridae" = "#212121",
             "Sofieviridae" = "#212121",
             "Claraviridae" = "#212121",
             "Pippiviridae" = "#4e342e", #VOC8
             "Siljaviridae" = "#4e342e",
             "Adrianviridae" = "#827717", #VOC4
             "Maisaviridae"  = "#455a64", #VOC2
             "Inaviridae" = "#455a64",
             "Arthurviridae" = "#455a64",
             "Mateoviridae" = "#455a64",
             "Solviridae" = "#455a64",
             "Ebbeviridae" = "#26a69a", #VOC3
             "Damienviridae" = "#f37735", #VOC5
             "Rasmusviridae" = "#d84315", #crassvirales
             "Celinaviridae" = "#d84315")

col.grid <- c(col.host, col.vfc)
```

```{r}
# sorted from right to left: VFCs highest loading to lowest; hostGenus highest number of vOTUs infecting to lowest
pdf("figures/EDFIG7.pdf", width=12, height=10)
par(cex = 0.75, mar = c(0, 0, 0, 0))
circos.par(start.degree = 90)
chordDiagram(df, order = c(loadings_df_vfc |> arrange(loading) |> pull(family), filt |> arrange(n) |> pull(hostGenus)),
             big.gap = 30,
             grid.col = col.grid,
             transparency = 0.3,
             scale = FALSE,
             preAllocateTracks = 1,
             annotationTrack = "grid")

# now, the image with rotated labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")
  circos.text(mean(xlim), ylim[1] + .1, sector.name, facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5), col = "black")
}, bg.border = NA)
circos.clear()
dev.off()
```

## Most prevalent VFC loadings and most abundant predicted host
```{r}
prevalence_df(physeqft_fam, taxa = T) |> 
  filter(tax %in% loadings_df_vfc$famid) |> 
  dplyr::select(-famid) |> rename(famid = tax) |> 
  # left_join(famid_most_abundant_host |> rename(famid = tax)) |> 
  left_join(family_host |> mutate(famid = paste0("f_", str_trim(famid))) |> dplyr::select(famid, host), by=c("famid")) |> 
  arrange(desc(prevalence)) |> 
  dplyr::select(family, famid, prevalence, host, everything()) |> 
  head()
```

## Less prevalent VFC loadings
```{r}
prevalence_df(physeqft_fam, taxa = T) |> 
  filter(tax %in% loadings_df_vfc$famid) |> 
  dplyr::select(-famid) |> rename(famid = tax) |> 
  left_join(family_host |> mutate(famid = paste0("f_", str_trim(famid))) |> dplyr::select(famid, host), by=c("famid")) |> 
  arrange(prevalence) |> 
  dplyr::select(family, famid, prevalence, host, everything()) |> 
  head()
```

## Strongest effect VFC loadings
```{r}
loadings_df_vfc |> 
  left_join(prevalence_df(physeqft_fam, taxa = T) |> 
              filter(tax %in% loadings_df_vfc$famid) |> 
              dplyr::select(-famid) |> rename(famid = tax), by="famid") |> 
  dplyr::select(family.x, famid, host, loading, everything()) |> 
  arrange(desc(abs(loading))) |> 
  head()
```

## FIG3C
```{r, fig.width=6, fig.height=4}
bac.ps <- phy.g |> 
  transform_phy(transform = "compositional") |> 
  transform_phy(transform = "log")
taxa_names(bac.ps) <- str_replace(taxa_names(bac.ps), "g_", "")
myfams <- loadings_df_vfc$famid
vir.ps <- subset_taxa(physeqf_fam, taxa_names(physeqf_fam) %in% myfams) |> # bacteriophage phyloseq object at the family level, not re-normalized to temperate
  subset_samples(abcno %in% sample_names(bac.ps)) |> 
  transform_phy("log") |> 
  microViz::ps_reorder(sample_names(bac.ps)) # reorder
taxa_names(vir.ps) <- tax_df(vir.ps)$family
vir <- otu_df(vir.ps)
bac <- otu_df(bac.ps)

cor.res <- t(cor(vir, bac, method = "spearman"))
cor.res2 <- as.data.frame(as.table(cor.res))

select_cols <- cor.res2 |>
  mutate(Var1 = as.character(Var1)) |>
  arrange(desc(Freq)) |>
  filter(Freq > 0.15 | Freq < -0.15) |> 
  pull(Var1) |> 
  unique()

pdf(file ="figures/FIG3C.pdf",
    width = 13, height = 9)
Heatmap(t(cor.res)[,select_cols],
        rect_gp = gpar(col = "white"),
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 10),
        column_names_rot = 45,
        column_dend_side = "bottom",
        column_names_side = "top",
        clustering_method_rows = "average",
        clustering_method_columns = "average",
        col=colorRamp2(c(-.2, 0, .4), c("navy", "white", "red")),
        heatmap_legend_param = list(title = "Spearman\n(rho)"),
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", t(cor.res)[,select_cols][i, j]), x, y, gp =
                      gpar(fontsize = 7, col="white"))})
dev.off()
```

## Save prediction
```{r}
save(splsda_model_relab_all_fam3, loadings_df_vfc,
   reps_auc, file="results/spls_vfc.RData")
```

# Virome scores
```{r, fig.width=6, fig.height=3}
medianRep <- splsda_model_relab_all_fam3 |>
  get_best_predictions() |> 
  group_by(Rep) |>
  summarize(auc = as.numeric(pROC::auc(obs, pred, direction = "<"))) |>
  arrange(desc(auc)) |>
  (\(x) x[5, "Rep"])() |> 
  as.character()

vir_score <- splsda_model_relab_all_fam3 |> 
  get_best_predictions() |> filter(Rep == medianRep) |> 
  pull(pred)

virome_pls_score_vfc <- data.frame(abcno = rownames(data),
                               pls_score = vir_score) |>
  left_join(dplyr::select(sample_df(physeqft), abcno, j45_5yr_ever, asthma_mother))

# ZSCORE of PLS score (scale and center)
virome_pls_score_vfc <- virome_pls_score_vfc |>
  mutate(zscore_vir_pls_score = as.numeric(scale(pls_score))) |> 
  rename(vir_pls_score = pls_score)
```

## virome scores vs. asthma
```{r,fig.width=4, fig.height=3}
glm(j45_5yr_ever ~ zscore_vir_pls_score,
    data=virome_pls_score_vfc, family="binomial") |>
  tidylog(exp=T, conf.int = T)
```

# 16S scores
```{r}
gut_scores <- sample_df(physeqf) |> 
  dplyr::select(abcno, j45_5yr_ever) |> 
  left_join(gut_score |>
              mutate(abcno = as.character(abcno))) |> 
  dplyr::rename(bac_pls_score = j45_cross_pls_1y_all) |> 
  mutate(zscore_bac_pls_score = as.numeric(scale(bac_pls_score))) |> 
  filter(!is.na(zscore_bac_pls_score))
```

```{r}
copdat_scores <- virome_pls_score_vfc |> 
  mutate(abcno = as.character(abcno)) |> 
  left_join(gut_scores |> mutate(abcno = as.character(abcno))) |> 
  filter(!is.na(bac_pls_score))
head(copdat_scores)
```

## 16S score vs. asthma
```{r}
glm(j45_5yr_ever ~ zscore_bac_pls_score,
    data=gut_scores, family="binomial") |>
  tidylog(exp=T, conf.int = T)
```

## mutually adjusted model
```{r}
glm(j45_5yr_ever ~ zscore_vir_pls_score + zscore_bac_pls_score, data=copdat_scores,
    family="binomial") |> 
  tidylog(exp=T, conf.int=T)
```

```{r}
write_tsv(virome_pls_score_vfc, file="results/virome_scores.tsv")
write_tsv(gut_scores, file="results/bac_scores.tsv")
write_tsv(copdat_scores, file="results/virome_bac_scores.tsv")
```

# EDFIG6
```{r}
myepi(copdat_scores, "zscore_vir_pls_score", "zscore_bac_pls_score", categorical = FALSE)
ggsave("figures/EDFIG6.pdf", device="pdf", dpi=300, height=2.5, width=3.5)
```

# session info
```{r}
sessionInfo()
```